D3064R0: How to Avoid OOTA Without Really Trying

### 2024-03-08 ootavacuous.tex
"Why rfe" appendix: Rework first paragraph and pull in JMM CT 4
Suggested-by: Alan Stern <stern@rowland.harvard.edu>
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

Although the examples presented in
Section~\ref{sec:OOTA: rf vs. rfe}
(and shown as litmus tests in
Listings~\ref{lst:OOTA With an Intra-Process Store-Load Link}
and~\ref{lst:OOTA With an Intra-Process Store-Load Link, Optimized})
demonstrate that rf can be atemporal, they demonstrate neither
OOTA cycles nor OOTA-like behavior in compiler-based C++ implementations.
This appendix provides more involved examples that can be optimized
so as to demonstrate OOTA-like behavior.

@@

@@

\subsection{Three-Process Version of JMM Causality Test Case 4}
\label{app:Three-Process Version of JMM Causality Test Case 4}

Listing~\ref{lst:Three-Process Version of JMM Causality Test Case 4}
expands
Listing~\ref{lst:Causality Test Case 4}
(discussed in
Appendix~\ref{app:Causality Test Case 4})
from two threads to three threads.
The \co{herd7} tool finds OOTA values, as expected.

All threads have straightforward semantic dependencies from their
relaxed atomic loads to their relaxed atomic stores, which prevents this
OOTA cycle on compiler-based C++ implementations.

>>>>

Listings~\ref{lst:OOTA With an Intra-Process Store-Load Link}
and~\ref{lst:OOTA With an Intra-Process Store-Load Link, Optimized}
show the examples from
Section~\ref{sec:OOTA: rf vs. rfe}
in the form of litmus tests.
A compiler is allowed to transform the program in
Listing~\ref{lst:OOTA With an Intra-Process Store-Load Link}
to the form of
Listing~\ref{lst:OOTA With an Intra-Process Store-Load Link, Optimized},
which demonstrates that intra-thread reads-from (rfi) links may be
eliminated by the compiler and therefore should not be used in a
definition of OOTA.
This section provides more involved examples to drive this point home.

Although the examples presented in
(and shown as litmus tests in
Listings~\ref{lst:OOTA With an Intra-Process Store-Load Link}
and~\ref{lst:OOTA With an Intra-Process Store-Load Link, Optimized})
demonstrate that rf can be atemporal, they demonstrate neither
OOTA cycles nor OOTA-like behavior in compiler-based C++ implementations.
This appendix provides more involved examples that can be optimized
so as to demonstrate OOTA-like behavior.

@@

The first step towards these more involved examples is shown in
Listing~\ref{lst:Three-Process Version of JMM Causality Test Case 4},
which expands
Listing~\ref{lst:Causality Test Case 4}
(discussed in
Appendix~\ref{app:Causality Test Case 4})
from two threads to three threads.
The \co{herd7} tool finds OOTA values, as expected.
All threads have straightforward semantic dependencies from their
relaxed atomic loads to their relaxed atomic stores, which prevents this
OOTA cycle on compiler-based C++ implementations.

@@
